A note on the instability and degeneracy of deep learning models

### 2016-12-02 writing/draft.Rmd
add formatting

---
title: "On the propriety of restricted Boltzmann machines"
author: | 
  | Andee Kaplan, Daniel Nordman, and Stephen Vardeman
  | Iowa State University
  | ajkaplan@iastate.edu, dnordman@iastate.edu, vardeman@iastate.edu
date: "December, 2016"
linestretch: 2 
output: 
  pdf_document:
    fig_caption: yes
    includes:
      in_header: latex/header.tex
    number_sections: yes
bibliography: references/refs.bib
abstract: |
  A restricted Boltzmann machine (RBM) is an undirected graphical model constructed for discrete or continuous random variables, with two layers, one hidden and one visible, and no conditional dependency within a layer. In recent years, RBMs have risen to prominence due to their connection to deep learning. By treating a hidden layer of one RBM as the visible layer in a second RBM, a deep architecture can be created. RBMs are thought to thereby have the ability to encode very complex and rich structures in data, making them attractive for supervised learning. However, the generative behavior of RBMs is largely unexplored. In this paper, we discuss the relationship between RBM parameter specification in the binary case and the tendency to undesirable model properties such as degeneracy, instability and uninterpretability. We also describe the difficulties that arise in likelihood-based and Bayes fitting of such (highly flexible) models, especially as Gibbs sampling (quasi-Bayes) methods are often advocated for the RBM model structure.
---

@@

A RBM model can be thought to exhibit an analogous form of *near-degeneracy* when there is a disproportionate amount of probability placed on a small number of elements in the sample space of visibles and hiddens, $\mathcal{C}^{\nv + \nh}$. Using the idea of @handcock2003assessing, when the random vector $\boldsymbol t(\boldsymbol x) = (v_1, \dots, v_\nv, h_1, \dots, h_\nh, v_1 h_1, \dots, v_V h_\nh ) \in \mathcal{T} \equiv \{\boldsymbol t(\boldsymbol x): \boldsymbol x \in \mathcal{C}^{\nh + \nv} \}$ from (\ref{eqn:t}), appearing in the neg-potential function $Q_{\boldsymbol \theta}(\cdot)$, has a mean vector $\boldsymbol \mu(\boldsymbol \theta) \in \mathbb{R}^{\nv + \nh + \nv * \nh}$ close to the boundary of the convex hull of $\mathcal{T}$, then the RBM model can be said to exhibit near-degeneracy at $\boldsymbol \theta \in \mathbb{R}^{\nv + \nh + \nh*\nv}$. Here the mean of $\boldsymbol t(\boldsymbol x)$ is
$$
\boldsymbol \mu(\boldsymbol \theta) = \text{E}_{\boldsymbol \theta} \boldsymbol t(\boldsymbol X)  = \sum\limits_{x \in \mathcal{C}^{\nv + \nh}} \left\{ \boldsymbol t(\boldsymbol x) f_{\boldsymbol \theta} (\boldsymbol x) \right\} = \sum\limits_{x \in \mathcal{C}^{\nv + \nh}} \left\{ \boldsymbol t(\boldsymbol x)\dfrac{\exp\left(\sum\limits_{i = 1}^\nv \sum\limits_{j=1}^\nh \theta_{ij} v_i h_j + \sum\limits_{i = 1}^\nv\theta_{v_i} v_i + \sum\limits_{j = 1}^\nh\theta_{h_j} h_j\right)}{\sum\limits_{\boldsymbol x \in \mathcal{C}^{\nh + \nv}}\exp\left(\sum\limits_{i = 1}^\nv \sum\limits_{j=1}^\nh \theta_{ij} v_i h_j + \sum\limits_{i = 1}^\nv\theta_{v_i} v_i + \sum\limits_{j = 1}^\nh\theta_{h_j} h_j\right)}\right\}.
$$

@@

S-unstable RBM model sequences are undesirable for several reasons. One is that, as mentioned above, small changes in data can lead to overly-sensitive changes in probability. Consider, for example, 
$$
\Delta(\boldsymbol \theta_\nv) \equiv \max \left\{\log \frac{P_{\boldsymbol \theta_\nv}(v_1, \dots, v_\nv)}{P_{\boldsymbol \theta_\nv}(v_1^*, \dots, v_\nv^*)} : (v_1, \dots, v_\nv) \& (v_1^*, \dots, v_\nv^*) \in \mathcal{C}^\nv \text{ differ in exactly one component}\right\},
$$
denoting the biggest log-probability ratio for a one component change in data outcomes (visibles) at a RBM parameter $\boldsymbol \theta_\nv$. We then have the following result.

@@

We can compute the mean vector for $\boldsymbol t(\boldsymbol x)$ as a function of the model parameters as 
\begin{align}
\boldsymbol \mu(\boldsymbol \theta) &= \text{E}_{\boldsymbol \theta}\left[ \boldsymbol t(\boldsymbol X) \right] =  \sum\limits_{\boldsymbol x = (v_1, h_1) \in \mathcal{X} = \{0,1\}^2} \left\{ t(x) \frac{\exp\left( \theta_{11} h_1 v_1 + \theta_{h1} h_1 + \theta_{v1} v_1\right)}{\gamma(\boldsymbol \theta)} \right\}
= \left[
\begin{matrix}
\frac{\exp\left(\theta_{v_1}\right) + \exp\left(\theta_{11} + \theta_{v_1} + \theta_{h_1}\right)}{\gamma(\boldsymbol \theta)} \\
\frac{\exp\left(\theta_{h_1}\right) + \exp\left(\theta_{11} + \theta_{v_1} + \theta_{h_1}\right)}{\gamma(\boldsymbol \theta)} \\
\frac{\exp\left(\theta_{11} + \theta_{v_1} + \theta_{h_1}\right)}{\gamma(\boldsymbol \theta)} \\
\end{matrix} \right]
\end{align}
where $\gamma(\boldsymbol \theta) = \sum\limits_{h_1 = 0}^1\sum\limits_{v_1 = 0}^1 \exp(\theta_{11} h_1 v_1 + \theta_{h_1}h_1 + \theta_{v_1}v_1)$. The three parametric coordinate functions of $\boldsymbol \mu(\boldsymbol \theta)$ can be represented as in Figure \ref{fig:degen_toy}. (Contour plots for three coordinate functions are shown in columns for various values of $\theta_{11}$.) In examining these, we see that as coordinates of $\boldsymbol \theta$ grow larger in magnitude, at least one mean function for the entries of $\boldsymbol t(\boldsymbol x)$ approaches a value 0 or 1, forcing $\boldsymbol \mu(\boldsymbol \theta) = \text{E}_{\boldsymbol \theta} \boldsymbol t(\boldsymbol x)$ to be near to the boundary of the convex hull of $\mathcal{T}$, as a sign of model near-degeneracy. Thus, for a very small example we can see the relationship between values of $\boldsymbol \theta$ and model degeneracy.

@@

marginal_acfs %>% ungroup() %>% select(-image_id) %>% rename(marginal = acf) %>%
  left_join(trunc_acfs %>% rename(truncated = acf)) %>%
  ggplot() +
  geom_segment(aes(x=lag, xend=lag, y=0, yend=truncated), colour = "black") +
  geom_segment(aes(x=lag, xend=lag, y=0, yend=marginal), colour = "red") +
  facet_wrap(~image_id) +
  xlab("Lag") +
  ylab("ACF")

>>>>

---
title: "On the propriety of restricted Boltzmann machines"
author: 
  - name: Andee Kaplan
    university: Iowa State University
    department: Department of Statistics
  - name: Daniel Nordman
    university: Iowa State University
    department: Department of Statistics
  - name: Stephen Vardeman
    university: Iowa State University
    department: Department of Statistics
output: 
  pdf_document:
    fig_caption: yes
    includes:
      in_header: latex/header.tex
    number_sections: yes
    template: latex/technometrics_article.tex
fontsize: 12pt
bibliography: references/refs.bib
abstract: |
  A restricted Boltzmann machine (RBM) is an undirected graphical model constructed for discrete or continuous random variables, with two layers, one hidden and one visible, and no conditional dependency within a layer. In recent years, RBMs have risen to prominence due to their connection to deep learning. By treating a hidden layer of one RBM as the visible layer in a second RBM, a deep architecture can be created. RBMs are thought to thereby have the ability to encode very complex and rich structures in data, making them attractive for supervised learning. However, the generative behavior of RBMs is largely unexplored. In this paper, we discuss the relationship between RBM parameter specification in the binary case and the tendency to undesirable model properties such as degeneracy, instability and uninterpretability. We also describe the difficulties that arise in likelihood-based and Bayes fitting of such (highly flexible) models, especially as Gibbs sampling (quasi-Bayes) methods are often advocated for the RBM model structure.
keywords: Degeneracy, Instability, Classification, Deep Learning, Graphical Models
blind: false
---

@@

A RBM model can be thought to exhibit an analogous form of *near-degeneracy* when there is a disproportionate amount of probability placed on a small number of elements in the sample space of visibles and hiddens, $\mathcal{C}^{\nv + \nh}$. Using the idea of @handcock2003assessing, when the random vector $\boldsymbol t(\boldsymbol x) = (v_1, \dots, v_\nv, h_1, \dots, h_\nh, v_1 h_1, \dots, v_V h_\nh ) \in \mathcal{T} \equiv \{\boldsymbol t(\boldsymbol x): \boldsymbol x \in \mathcal{C}^{\nh + \nv} \}$ from (\ref{eqn:t}), appearing in the neg-potential function $Q_{\boldsymbol \theta}(\cdot)$, has a mean vector $\boldsymbol \mu(\boldsymbol \theta) \in \mathbb{R}^{\nv + \nh + \nv * \nh}$ close to the boundary of the convex hull of $\mathcal{T}$, then the RBM model can be said to exhibit near-degeneracy at $\boldsymbol \theta \in \mathbb{R}^{\nv + \nh + \nh*\nv}$. Here the mean of $\boldsymbol t(\boldsymbol x)$ is
\begin{align*}
\boldsymbol \mu(\boldsymbol \theta) = \text{E}_{\boldsymbol \theta} \boldsymbol t(\boldsymbol X)  &= \sum\limits_{x \in \mathcal{C}^{\nv + \nh}} \left\{ \boldsymbol t(\boldsymbol x) f_{\boldsymbol \theta} (\boldsymbol x) \right\} \\
&= \sum\limits_{x \in \mathcal{C}^{\nv + \nh}} \left\{ \boldsymbol t(\boldsymbol x)\dfrac{\exp\left(\sum\limits_{i = 1}^\nv \sum\limits_{j=1}^\nh \theta_{ij} v_i h_j + \sum\limits_{i = 1}^\nv\theta_{v_i} v_i + \sum\limits_{j = 1}^\nh\theta_{h_j} h_j\right)}{\sum\limits_{\boldsymbol x \in \mathcal{C}^{\nh + \nv}}\exp\left(\sum\limits_{i = 1}^\nv \sum\limits_{j=1}^\nh \theta_{ij} v_i h_j + \sum\limits_{i = 1}^\nv\theta_{v_i} v_i + \sum\limits_{j = 1}^\nh\theta_{h_j} h_j\right)}\right\}.
\end{align*}

@@

S-unstable RBM model sequences are undesirable for several reasons. One is that, as mentioned above, small changes in data can lead to overly-sensitive changes in probability. Consider, for example, 
$$
\Delta(\boldsymbol \theta_\nv) \equiv \max \left\{\log \frac{P_{\boldsymbol \theta_\nv}(\boldsymbol v)}{P_{\boldsymbol \theta_\nv}(\boldsymbol v^*)} : \boldsymbol v \text{ }\& \text{ } \boldsymbol v^* \in \mathcal{C}^\nv \text{ differ in exactly one component}\right\},
$$
denoting the biggest log-probability ratio for a one component change in data outcomes (visibles) at a RBM parameter $\boldsymbol \theta_\nv$. We then have the following result.

@@

We can compute the mean vector for $\boldsymbol t(\boldsymbol x)$ as a function of the model parameters as 
\begin{align*}
\boldsymbol \mu(\boldsymbol \theta) = \text{E}_{\boldsymbol \theta}\left[ \boldsymbol t(\boldsymbol X) \right] =  \sum\limits_{\boldsymbol x = (v_1, h_1) \in \{0,1\}^2} \left\{ t(x) \frac{\exp\left( \theta_{11} h_1 v_1 + \theta_{h1} h_1 + \theta_{v1} v_1\right)}{\gamma(\boldsymbol \theta)} \right\} = \left[
\begin{matrix}
\frac{\exp\left(\theta_{v_1}\right) + \exp\left(\theta_{11} + \theta_{v_1} + \theta_{h_1}\right)}{\gamma(\boldsymbol \theta)} \\
\frac{\exp\left(\theta_{h_1}\right) + \exp\left(\theta_{11} + \theta_{v_1} + \theta_{h_1}\right)}{\gamma(\boldsymbol \theta)} \\
\frac{\exp\left(\theta_{11} + \theta_{v_1} + \theta_{h_1}\right)}{\gamma(\boldsymbol \theta)} \\
\end{matrix} \right]
\end{align*}
where $\gamma(\boldsymbol \theta) = \sum\limits_{h_1 = 0}^1\sum\limits_{v_1 = 0}^1 \exp(\theta_{11} h_1 v_1 + \theta_{h_1}h_1 + \theta_{v_1}v_1)$. The three parametric coordinate functions of $\boldsymbol \mu(\boldsymbol \theta)$ can be represented as in Figure \ref{fig:degen_toy}. (Contour plots for three coordinate functions are shown in columns for various values of $\theta_{11}$.) In examining these, we see that as coordinates of $\boldsymbol \theta$ grow larger in magnitude, at least one mean function for the entries of $\boldsymbol t(\boldsymbol x)$ approaches a value 0 or 1, forcing $\boldsymbol \mu(\boldsymbol \theta) = \text{E}_{\boldsymbol \theta} \boldsymbol t(\boldsymbol x)$ to be near to the boundary of the convex hull of $\mathcal{T}$, as a sign of model near-degeneracy. Thus, for a very small example we can see the relationship between values of $\boldsymbol \theta$ and model degeneracy.

@@

marginal_acfs %>% ungroup() %>% select(-image_id) %>% rename(marginal = acf) %>%
  left_join(trunc_acfs %>% rename(truncated = acf)) %>%
  ggplot() +
  geom_segment(aes(x=lag, xend=lag, y=0, yend=truncated), colour = "black") +
  geom_segment(aes(x=lag, xend=lag, y=0, yend=marginal), colour = "red") +
  geom_point(aes(x=lag, y=marginal), colour = "red", size = 0.5) +
  facet_wrap(~image_id) +
  xlab("Lag") +
  ylab("ACF")

### 2016-12-02 writing/note.Rmd
add formatting

---
title: "A note on the instability and degeneracy of deep learning models"
author: | 
  | Andee Kaplan, Daniel Nordman, and Stephen Vardeman
  | Iowa State University
  | ajkaplan@iastate.edu, dnordman@iastate.edu, vardeman@iastate.edu
date: "December, 2016"
linestretch: 2
output: 
  pdf_document:
    keep_tex: true
    fig_caption: yes
    includes:
      in_header: latex/header.tex
    number_sections: yes
bibliography: references/refs.bib
abstract: |
  A probability model exhibits instability if small changes in a data outcome result in large, and often unanticipated, changes in probability. For correlated data structures found in several application areas, there is increasing interest in predicting/identifying instability. We consider the problem of quantifying instability for general probability models defined on sequences of observations, where each sequence of length $N$ has a finite number of possible outcomes. (A sequence of probability models results indexed by $N$ that accommodates data of expanding dimension.) Model instability is formally shown to occur when a certain log-probability ratio under such models grows faster than $N$. In this case, a one component change in the data sequence can shift probability by orders of magnitude. Also, as a measure of instability becomes more extreme, the resulting probability models are shown  to tend to degeneracy, placing all their probability on arbitrarily small portions of the sample space. These results on instability apply to large classes of models commonly used in random graphs, network analysis, and machine learning contexts.
---

@@

**Deep Boltzmann machine (DBM).** The DBM class of models maintains conditional independence within all layers in the model by stacking RBM models and maintaining the restriction of only allowing conditional dependence between neighboring layers. The joint probability mass function for a DBM is
$$
P_{\boldsymbol \theta_N} (\tilde{\boldsymbol x}) = \exp\left[ \sum\limits_{i = 1}^M\boldsymbol \alpha^{(i)T} \boldsymbol h^{(i)} + \boldsymbol \beta^T \boldsymbol x + \boldsymbol h^{(1)T} \Gamma^{(0)} \boldsymbol x + \sum\limits_{i = 1}^{M - 1} \boldsymbol h^{(i)T} \Gamma^{(i)} \boldsymbol h^{(i + 1)} - \psi(\boldsymbol \theta_N) \right], 
$$
for $\tilde{\boldsymbol x} = (\boldsymbol h^{(1)}, \dots, \boldsymbol h^{(M)}, \boldsymbol x) \in \mathcal{X}^{N_{(H,1)} + \cdots + N_{(H,M)} +N}$ where
$$
\psi(\boldsymbol \theta_N) = \log \sum\limits_{\tilde{\boldsymbol x} \in \mathcal{X}^{N_{(H,1)} + \cdots + N_{(H,M)} +N}} \exp\left[ \sum\limits_{i = 1}^M\boldsymbol \alpha^{(i)T} \boldsymbol h^{(i)} + \boldsymbol \beta^T \boldsymbol x + \boldsymbol h^{(1)T} \Gamma^{(0)} \boldsymbol x + \sum\limits_{i = 1}^{M - 1} \boldsymbol h^{(i)T} \Gamma^{(i)} \boldsymbol h^{(i + 1)}\right], \qquad \boldsymbol \theta_N \in \Theta_N
$$ 
is the normalizing function and parameters in the model are $\boldsymbol \beta \in \mathbb{R}^V$, $\boldsymbol \alpha^{(i)} \in \mathbb{R}^{N_{(H,i)}}$ for $i = 1, \dots, M$, along with a matrix $\Gamma^{(0)}$ of dimension $N_{(H,1)} \times N$, and matrices $\Gamma^{(i)}$ of dimension $N_{(H,i)} \times N_{(H,i+1)}$ for $i = 1, \dots, M-1$. Let $\boldsymbol \theta_N = (\boldsymbol \alpha^{(1)}, \dots, \boldsymbol \alpha^{(M)}, \boldsymbol \beta,\Gamma^{(0)}, \dots, \Gamma^{(M - 1)}) \in \Theta_N \subset \mathbb{R}^{q(N)}$ denote the combined vector of parameters with total length $q(N)= N_{(H,1)}+\cdots N_{(H,M)} + N + N_{(H,1)}*N+N_{H,2}*H_{(H,1)}+\cdots +N_{(H,M)}*H_{(H,M)-1}$.

@@

S-unstable FSFS model sequences are undesirable for several reasons. One is that small changes in data can lead to overly-sensitive changes in probability. Consider, for example, 
$$
\Delta(\boldsymbol \theta_N) \equiv \max \left\{\log \frac{P_{\boldsymbol \theta_N}(x_1, \dots, x_N)}{P_{\boldsymbol \theta_N}(x_1^*, \dots, x_N^*)} : (x_1, \dots, x_N) \& (x_1^*, \dots, x_N^*) \in \mathcal{X}^N \text{ differ in exactly one component}\right\},
$$
the biggest log-probability ratio for a one component change in data outcomes at a FSFS parameter $\boldsymbol \theta_N$. We then have the following (non-asymptotic) result.

@@

\begin{remark}
There is a further generalization the notion of instability in Definition \ref{defn:instab} meant to address independent replications of data sequences. That is, one might consider data as $n$ independent and identically distributed replications $\boldsymbol X_1, \dots, \boldsymbol X_n$, where each $\boldsymbol X_i=(X_{i,1},\ldots,X_{i,N}) \in \mathcal{X}^N$ follows a common FSFS model with probabilities $P_{\theta_N}(\boldsymbol x)>0$,  $\boldsymbol x\in\mathcal{X}^N$ and $|\mathcal{X}|<\infty$, for $i=1,\ldots,n$.  This leads to a total of $n*N$ random variables in the joint model. However, the definition of S-unstable and Propositions \ref{prop:instab}-\ref{prop:degen} still hold for such iid replications. This is because $\left(\max\limits_{\boldsymbol x \in \mathcal{X}^N}P_{\boldsymbol \theta_N}(\boldsymbol x)\right)^n$ is the largest probability possible under the joint model for the $n$ replications while $\left(\min\limits_{\boldsymbol x \in \mathcal{X}}P_{\boldsymbol \theta_N}(\boldsymbol x)\right)^n$ is the smallest probability. Thus, for the combined replications $\boldsymbol X_1, \ldots, \boldsymbol X_n$, the analog definition of the extremal log-probability becomes
$$
\frac{\text{extremal log-probability ratio}}{\text{\# random variables in the model}} \equiv \frac{1}{n*N}\log \left[\frac{\left(\max\limits_{\boldsymbol x \in \mathcal{X}^N}P_{\boldsymbol \theta_N}(\boldsymbol x)\right)^n}{\left(\min\limits_{\boldsymbol x \in \mathcal{X}^N}P_{\boldsymbol \theta_N}(\boldsymbol x)\right)^n}\right] = \frac{1}{N} \log \left[\frac{\max\limits_{\boldsymbol x \in \mathcal{X}^N}P_{\boldsymbol \theta_N}(\boldsymbol x)}{\min\limits_{\boldsymbol x \in \mathcal{X}^N}P_{\boldsymbol \theta_N}(\boldsymbol x)}\right] = \frac{1}{N} \text{ELPR}(\boldsymbol \theta_N),
$$
implying that the definition of an S-unstable FSFS model sequence is invariant to the level ($n$) of independent replication. Consequently, overall model instabilities may be characterized by those of one observation from the common FSFS model.   
\end{remark}

>>>>

---
title: "A note on the instability and degeneracy of deep learning models"
author: 
  - name: Andee Kaplan
    affiliation: Iowa State University
    department: Department of Statistics
  - name: Daniel Nordman
    affiliation: Iowa State University
    department: Department of Statistics
  - name: Stephen Vardeman
    affiliation: Iowa State University
    department: Department of Statistics
output: 
  pdf_document:
    keep_tex: true
    fig_caption: yes
    includes:
      in_header: latex/header.tex
    number_sections: yes
    template: latex/technometrics_article.tex
fontsize: 12pt
bibliography: references/refs.bib
abstract: |
  A probability model exhibits instability if small changes in a data outcome result in large, and often unanticipated, changes in probability. For correlated data structures found in several application areas, there is increasing interest in predicting/identifying instability. We consider the problem of quantifying instability for general probability models defined on sequences of observations, where each sequence of length $N$ has a finite number of possible outcomes. (A sequence of probability models results indexed by $N$ that accommodates data of expanding dimension.) Model instability is formally shown to occur when a certain log-probability ratio under such models grows faster than $N$. In this case, a one component change in the data sequence can shift probability by orders of magnitude. Also, as a measure of instability becomes more extreme, the resulting probability models are shown  to tend to degeneracy, placing all their probability on arbitrarily small portions of the sample space. These results on instability apply to large classes of models commonly used in random graphs, network analysis, and machine learning contexts.
keywords: Degeneracy, Instability, Deep Learning, Graphical Models
blind: false
---

@@

**Deep Boltzmann machine (DBM).** The DBM class of models maintains conditional independence within all layers in the model by stacking RBM models and maintaining the restriction of only allowing conditional dependence between neighboring layers. The joint probability mass function for a DBM is
$$
P_{\boldsymbol \theta_N} (\tilde{\boldsymbol x}) = \exp\left[ \sum\limits_{i = 1}^M\boldsymbol \alpha^{(i)T} \boldsymbol h^{(i)} + \boldsymbol \beta^T \boldsymbol x + \boldsymbol h^{(1)T} \Gamma^{(0)} \boldsymbol x + \sum\limits_{i = 1}^{M - 1} \boldsymbol h^{(i)T} \Gamma^{(i)} \boldsymbol h^{(i + 1)} - \psi(\boldsymbol \theta_N) \right], 
$$
for $\tilde{\boldsymbol x} = (\boldsymbol h^{(1)}, \dots, \boldsymbol h^{(M)}, \boldsymbol x) \in \mathcal{X}^{N_{(H,1)} + \cdots + N_{(H,M)} +N}$ where
$$
\psi(\boldsymbol \theta_N) = \log \sum\limits_{\tilde{\boldsymbol x} \in \mathcal{X}^{N_{(H,1)} + \cdots + N_{(H,M)} +N}} \exp\left[ \sum\limits_{i = 1}^M\boldsymbol \alpha^{(i)T} \boldsymbol h^{(i)} + \boldsymbol \beta^T \boldsymbol x + \boldsymbol h^{(1)T} \Gamma^{(0)} \boldsymbol x + \sum\limits_{i = 1}^{M - 1} \boldsymbol h^{(i)T} \Gamma^{(i)} \boldsymbol h^{(i + 1)}\right],
$$ 
for $\boldsymbol \theta_N \in \Theta_N$ is the normalizing function and parameters in the model are $\boldsymbol \beta \in \mathbb{R}^V$, $\boldsymbol \alpha^{(i)} \in \mathbb{R}^{N_{(H,i)}}$ for $i = 1, \dots, M$, along with a matrix $\Gamma^{(0)}$ of dimension $N_{(H,1)} \times N$, and matrices $\Gamma^{(i)}$ of dimension $N_{(H,i)} \times N_{(H,i+1)}$ for $i = 1, \dots, M-1$. Let $\boldsymbol \theta_N = (\boldsymbol \alpha^{(1)}, \dots, \boldsymbol \alpha^{(M)}, \boldsymbol \beta,\Gamma^{(0)}, \dots, \Gamma^{(M - 1)}) \in \Theta_N \subset \mathbb{R}^{q(N)}$ denote the combined vector of parameters with total length $q(N)= N_{(H,1)}+\cdots N_{(H,M)} + N + N_{(H,1)}*N+N_{H,2}*H_{(H,1)}+\cdots +N_{(H,M)}*H_{(H,M)-1}$.

@@

S-unstable FSFS model sequences are undesirable for several reasons. One is that small changes in data can lead to overly-sensitive changes in probability. Consider, for example, 
$$
\Delta(\boldsymbol \theta_N) \equiv \max \left\{\log \frac{P_{\boldsymbol \theta_N}(\boldsymbol x)}{P_{\boldsymbol \theta_N}(\boldsymbol x^*)} : \boldsymbol x \text{ }\& \text{ } \boldsymbol x^* \in \mathcal{X}^N \text{ differ in exactly one component}\right\},
$$
the biggest log-probability ratio for a one component change in data outcomes at a FSFS parameter $\boldsymbol \theta_N$. We then have the following (non-asymptotic) result.

@@

\begin{remark}
There is a further generalization the notion of instability in Definition \ref{defn:instab} meant to address independent replications of data sequences. That is, one might consider data as $n$ independent and identically distributed replications $\boldsymbol X_1, \dots, \boldsymbol X_n$, where each $\boldsymbol X_i=(X_{i,1},\ldots,X_{i,N}) \in \mathcal{X}^N$ follows a common FSFS model with probabilities $P_{\theta_N}(\boldsymbol x)>0$,  $\boldsymbol x\in\mathcal{X}^N$ and $|\mathcal{X}|<\infty$, for $i=1,\ldots,n$.  This leads to a total of $n*N$ random variables in the joint model. However, the definition of S-unstable and Propositions \ref{prop:instab}-\ref{prop:degen} still hold for such iid replications. This is because $\left(\max\limits_{\boldsymbol x \in \mathcal{X}^N}P_{\boldsymbol \theta_N}(\boldsymbol x)\right)^n$ is the largest probability possible under the joint model for the $n$ replications while $\left(\min\limits_{\boldsymbol x \in \mathcal{X}}P_{\boldsymbol \theta_N}(\boldsymbol x)\right)^n$ is the smallest probability. Thus, for the combined replications $\boldsymbol X_1, \ldots, \boldsymbol X_n$, the analog definition of the extremal log-probability becomes
\begin{align*}
\frac{\text{extremal log-probability ratio}}{\text{\# random variables in the model}} &\equiv \frac{1}{n*N}\log \left[\frac{\left(\max\limits_{\boldsymbol x \in \mathcal{X}^N}P_{\boldsymbol \theta_N}(\boldsymbol x)\right)^n}{\left(\min\limits_{\boldsymbol x \in \mathcal{X}^N}P_{\boldsymbol \theta_N}(\boldsymbol x)\right)^n}\right] = \frac{1}{N} \log \left[\frac{\max\limits_{\boldsymbol x \in \mathcal{X}^N}P_{\boldsymbol \theta_N}(\boldsymbol x)}{\min\limits_{\boldsymbol x \in \mathcal{X}^N}P_{\boldsymbol \theta_N}(\boldsymbol x)}\right] \\
&= \frac{1}{N} \text{ELPR}(\boldsymbol \theta_N),
\end{align*}
implying that the definition of an S-unstable FSFS model sequence is invariant to the level ($n$) of independent replication. Consequently, overall model instabilities may be characterized by those of one observation from the common FSFS model.   
\end{remark}
